A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines

### 2019-07-29 latex/paper.tex
Edits in results

In the past few years, neuroimaging has entered the Big Data era due to the joint
increase in image resolution, data sharing, and study sizes. However, no particular
Big Data engines has emerged in this field, and several alternatives remain
available. We compare two popular Big Data engines with Python APIs, Apache Spark and
Dask, for their runtime performance in processing neuroimaging pipelines. Our
evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} Big Brain
image, and a real pipeline processing anatomical data from more than 1,000 subjects.
We benchmark these pipelines using various combinations of task durations, data
sizes, and numbers of workers, deployed on a 8-node (8 cores ea.) compute cluster in
Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
Delayed and Futures. Results show that despite slight differences between Spark and
Dask, both frameworks perform comparably. However, Dask pipelines risk being limited
by Python's GIL depending on task type and cluster configuration. In all cases, the
major limiting factor was data transfer. While either framework is suitable for
neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
\end{abstract}

@@

First and foremost, Spark is written in Scala while Dask is in Python.
Given the popularity of Python in scientific communities, this arguably
gives an edge to Dask due to data serialization costs from Python to Scala.
On the other hand, Python's Global Interpreter Lock (GIL) might reduce
parallelism in some cases. This difference in programming languages also
has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
ecosystem, Dask provides almost transparent parallelization of applications
manipulating Numpy arrays or Pandas data frames. On the other hand, Spark's
Java, R and Python APIs allow to easily parallelize analyses that combine
these languages, with reduced performance loss. Our study focuses on
performance, although we recognize that other factors will also impact the
choice of a framework in practice.

Spark and Dask were both included in the evaluation reported
in~\cite{Mehta:17}, where a neuroimaging application
processed approximately \SI{100}{\giga\byte} of data. In this work, Dask was
reported to have a slight performance advantage over Spark. Overall, Dask's
end-to-end time (makespan) was measured to be up to 14\% faster than Spark,
due to ``more efficient pipelining'' %(Fig 10) 
and serialization time to
Python. % (Fig. 12-a)
Dask, however, was reported to have a larger startup
time than Spark. % (Fig. 12-c)
The analysis remained at a quite high level
though, leaving most of the observed performance difference unexplained. In
comparison, our study will provide a detailed analysis of performance
differences and similarities between Spark and Dask for neuroimaging data processing.

@@

Figure~\ref{fig:inc_tt_worker} shows the total execution time of the
Incrementation application, broken down into data transfer (read and
write), compute (sleep), and overhead time. As expected, the computing time
stays similar when the number of workers increases. However, the data
transfer time and overhead increase proportionally to the number of
workers with regression slopes: Spark (\SI{2337}{\second/task}), Bag
(\SI{2650}{\second/task}), Delayed (\SI{3251}{\second/task}), Futures
(\SI{3570}{\second/task}).

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's, particularly when the number of workers increase.
Within Dask, Delayed and Futures have a higher overhead than Bag. However,
overhead differences are compensated by an increase in data transfer time,
as a reduced overhead increases the concurrency between data transfers.

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
number of image blocks for constant BigBrain image size. We were not able to run
Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size it can
compute. Once again, we do not observe any significant difference among the engines.
For all engines, makespan variability increases with the number of blocks, however,
engines scale very well in general.

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks, Delayed and Futures have much higher overhead. This is because they are
thread-based and use multiple threads to compute a block which affect the calculation
of the overhead. Once again, the data transfer time reduces with more blocks however
the overhead time increases by a similar amount. This is not observed for 30 blocks
as the workers are not used at full capacity, i.e., some threads are idle. Finally,
the variability of the overhead increases with the number of blocks, which explains
the makespan variability mentioned previously.

@@

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall, Spark and Dask APIs are once
again equivalent, although Delayed and Futures are slightly faster
than Bag and, Spark for 1 and 10 iterations, and Futures are faster than
Delayed, Bag and Spark for 100 iterations. Differences remain minor
though.

@@

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine. Gantt
charts are structured in batches of up to 64 read-compute-write concurrent sequences.
Files read in the first batch are much longer than the ones in the following batches:
this is due to the high synchronization of data transfers that leads to a high
saturation of the shared file system. We also note that overhead, represented in
white, is concentrated around the data transfer tasks, and the
computing tasks that run concurrently with data transfers.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/spark-inc-baseline.html?download=1}{
            \includegraphics[clip,width=\columnwidth,
            height=0.15\textheight]{images/spark_inc_gantt.png}}
        \caption{Spark}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/dask-bag-inc-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_inc_gantt.png}}
        \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/dask-delayed-inc-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_inc_gantt.png}}
        \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/dask-futures-inc-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/futures_inc_gantt.png}}
        \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{Incrementation: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
8 CPUs/worker}\label{fig:inc_gantt}
\end{figure}

%%% HISTOGRAM %%%
\subsection{Histogram: Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
amounts of workers. Spark is significantly faster than Dask APIs. The difference
narrows as the number of workers increases but stays significant. Between Dask APIs
there is no substantial difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
shown. The compute time is significantly larger for Dask than for Spark. This is
potentially due to Python's GIL preventing Dask to parallelize the computation
over
multiple threads. Overall, the IO and overhead are comparable for all engines.

On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
when their number of workers varies, except for 8 workers where it increases slightly.
On the other hand, Spark total execution time keeps increasing; especially at 8
workers. This is because Dask engines benefit more from additional workers, in
this GIL-bounded scenario.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
        \caption{Histogram total time}\label{fig:histo_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 8 CPUs/worker}
\end{figure}

\subsection{Histogram: Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block sizes. 
Once more, Spark is significantly faster than Dask engines and Bag is slightly
faster than Delayed. Overall, the engines do not react to changes in block size.

In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
Note that at 30 blocks the workers' resources are not fully used since the
workers can process up to 64 tasks in parallel. Moreover, the overhead for
30 blocks is erroneous due to a lower amount of block than available threads.
This leads to thread based engines to have higher overhead. Given that, 
Spark has a much lower compute time which makes it significantly faster
than Dask engines. Overall, for all engines, lower block size results in lower
IO time but increases the overhead time. Again, this is due to large overhead
desynchronizing task thus reducing the NFS bottleneck.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
        \caption{Histogram makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
        \caption{Histogram total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers, 8 CPUs/worker}
\end{figure}
% \TG{We decided to omit Dask Futures as it does not have any value over Dask
% Delayed in this application.}

\subsection{Histogram: Gantt chart}
Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline Histogram
experiment. Spark overhead is distributed through all the execution while Bag
overhead happens mostly when a thread reads subsequent blocks and
Delayed overhead is concentrated after reading all the blocks.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/spark-histo-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/spark_histo_gantt.png}}
        \caption{Spark}\label{fig:histo_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/dask-bag-histo-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_histo_gantt.png}}
        \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_histo_gantt.png}}
        \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
    \end{subfigure}
    \caption{Histogram: 125 blocks, 8 workers, 8
    CPUs/worker}\label{fig:histo_gantt}
\end{figure}

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
Overall, the makespan for all engines does not change considerably with an increase
in the number of workers. This is likely a result of using a non-distributed
shared file system. There is no
substantial difference between all engines, although, Delayed is slightly faster.

From Figure~\ref{fig:histo_np_tt_worker}, the total execution time breakdown is
shown. The overhead and read time increase proportionally to the number of
workers thus increasing the total time. This is because (1) most workers read
at the same time due to a low computation-IO ratio and (2) there is more
intra-worker and inter-worker communication.

@@

In figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all
engines. However, when computation starts the subsequent read tasks are much
shorter for Delayed. This is due the desynchronization of the task reducing the
IO bottleneck.

\begin{figure}[!tb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/spark-histo_np-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/spark_histo_np_gantt.png}}
        \caption{Spark}\label{fig:histo_np_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/dask-bag-histo_np-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_histo_np_gantt.png}}
        \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo_np-baseline.html?download=1}{
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_histo_np_gantt.png}}
        \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
    \end{subfigure}
    \caption{Histogram (NumPy): 125 blocks, 8 workers, 8
    CPUs/worker}\label{fig:histo_np_gantt}
\end{figure}

%%% BIDS EXAMPLE %%%
\subsection{BidsApp example: Number of workers}
Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when
varying the amount of workers. Overall, there is no substantial difference
between the engines. The makespan scales well with the increase in workers,
however, the scaling reduces considerably when reaching 8 workers. This is due
to an increase in overhead and an IO bottleneck caused parallel accesses to 
a non-distributed file system.

In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
shown. Each engine has a similar total time independently of its number of workers.
This is due to the, almost, linear scaling of the engines. The overhead increases
proportionally to the number of workers with regression slopes: Spark
(\SI{224}{\second/task}), Bag (\SI{183}{\second/task}), Delayed
(\SI{173}{\second/task}), Futures (\SI{132}{\second/task}). This is due to more
communication between the scheduler and workers as well as intra-worker and
inter-worker communication.

>>>>

In the past few years, neuroimaging has entered the Big Data era due to the
joint increase in image resolution, data sharing, and study sizes. However,
no particular Big Data framework has emerged in this field, and several
alternatives remain available. We compare two popular Big Data frameworks
with Python APIs, Apache Spark and Dask, for their runtime performance in
processing neuroimaging pipelines. Our evaluation uses two synthetic
pipelines processing the 81~GB Big Brain image, and a real pipeline
processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
 various combinations of task durations, data sizes, and numbers of workers,
deployed on a 64-node compute cluster in Compute Canada's Arbutus
cloud. We evaluate PySpark's RDD API against Dask's Bag, Delayed and Futures.
Results show that despite slight differences between Spark and Dask, both
frameworks perform comparably. However, Dask pipelines risk being limited
by Python's GIL depending on task type and cluster configuration. In all
cases, the major limiting factor was data transfer. While either framework
is suitable for neuroimaging pipelines, more effort needs to be placed in
reducing data transfer time.
\end{abstract}

@@

% Differences
First and foremost, Spark is written in Scala while Dask is in Python.
Given the popularity of Python in scientific communities, this arguably
gives an edge to Dask due to data serialization costs from Python to Scala.
On the other hand, Python's Global Interpreter Lock (GIL) might reduce
parallelism in some cases. This difference in programming languages also
has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
ecosystem, Dask provides almost transparent parallelization of applications
manipulating Numpy arrays or Pandas data frames. On the other hand, Spark's
Java, R and Python APIs allow to easily parallelize analyses that combine
these languages, with reduced performance loss. Our study focuses on
performance, although we recognize that other factors will also impact the
choice of a framework in practice.

% Related work
Spark and Dask were both included in the evaluation reported
in~\cite{Mehta:17}, where a neuroimaging application
processed approximately \SI{100}{\giga\byte} of data. In this work, Dask was
reported to have a slight performance advantage over Spark. Overall, Dask's
end-to-end time (makespan) was measured to be up to 14\% faster than Spark,
due to ``more efficient pipelining'' %(Fig 10) 
and serialization time to
Python. % (Fig. 12-a)
Dask, however, was reported to have a larger startup
time than Spark. % (Fig. 12-c)
The analysis remained at a quite high level
though, leaving most of the observed performance difference unexplained. In
comparison, our study will provide a detailed analysis of performance
differences and similarities between Spark and Dask for neuroimaging data processing.

@@

Figure~\ref{fig:inc_tt_worker} shows the total execution time of the
Incrementation application, broken down into data transfer (read and
write), compute (sleep), and overhead time. As expected, the computing time
stays similar when the number of workers increases. However, the data
transfer time and overhead increase substantially with the number of
workers.

% \hspace{1cm}\(\hat{y}=2336.9499\cdot{x} + 325.00574\) (Spark)

% \hspace{1cm}\(\hat{y}=2650.13505\cdot{x} + 1585.37218\) (Bag)

% \hspace{1cm}\(\hat{y}=3250.53406\cdot{x} + 1769.50351\) (Delayed)

% \hspace{1cm}\(\hat{y}=3570.39758\cdot{x} + 1062.23031\) (Futures)

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's, particularly when the number of workers
increase. Moreover, Delayed and Futures have a higher overhead than Bag.
However, overhead differences are compensated by an increase in data
transfer time, as a reduced overhead increases the concurrency between data
transfers and the application saturates the 10Gbits/s network bandwidth.

@@

Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
number of image blocks for constant BigBrain image size. We were not able to run
Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size.
 Once again, we do not observe any substantial difference among the engines.
For all engines, makespan variability increases with the number of blocks, however,
engines scale very well in general.

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks, the increased overhead of Delayed and Futures is a calculation artifact 
coming from the fact that only 30 of the 64 available threads can be used concurrently. Once again, the data transfer time reduces with more blocks but 
the overhead time increases by a similar amount. This is not observed for 30 blocks
as the workers are not used at full capacity, i.e., some threads are idle. Finally,
the variability of the overhead increases with the number of blocks, which explains
the makespan variability mentioned previously.

@@

%% Iterations
\subsection{Incrementation: Number of iterations}
Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
varying the number of iterations. Overall, Spark and Dask APIs are once
again equivalent, although Delayed and Futures are slightly faster
than Bag and Spark for 1 and 10 iterations, and Futures are faster than
Delayed, Bag and Spark for 100 iterations. Differences remain minor
though.

@@

Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine. Gantt
charts are structured in batches of up to 64 read-compute-write concurrent sequences.
File reads in the first batch are much longer than the ones in the following batches:
this is due to the high synchronization of data transfers that leads to a high
saturation of the shared file system. We also note that overhead, represented in
white, is concentrated around the data transfer tasks and the
computing tasks that run concurrently with data transfers.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/spark_inc_gantt.png}
        \caption{Spark}\label{fig:inc_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_inc_gantt.png}%
        \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_inc_gantt.png}%
        \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/futures_inc_gantt.png}%
        \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
    \end{subfigure}
    \caption{Incrementation: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
8 CPUs/worker}\label{fig:inc_gantt}
\end{figure}

%%% HISTOGRAM %%%
\subsection{Histogram (Python): Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the Histogram application for various
amounts of workers. Spark is significantly faster than Dask APIs. The difference
narrows as the number of workers increases but it remains significant. Between Dask APIs
there is no substantial difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
shown. The compute time is significantly larger for Dask than for Spark. This is
presumably due to Python's GIL preventing Dask to parallelize the computation
over
multiple threads. Overall, the I/O and overhead times are comparable for all engines.

Figure~\ref{fig:histo_tt_worker} shows that Dask engines have a similar
total execution time when their number of workers varies, except for 8
workers where it increases slightly due to increased read time.
On the other hand, Spark total execution time keeps increasing; especially at 8
workers. This is because Dask engines benefit more from additional workers, in
this GIL-bounded scenario.

\begin{figure}[!b]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
        \caption{Histogram (Python) makespan}\label{fig:histo_ms_worker}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
        \caption{Histogram (Python) total time}\label{fig:histo_tt_worker}
    \end{subfigure}
    \caption{125 blocks, 8 CPUs/worker}
\end{figure}

\subsection{Histogram (Python): Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block sizes. 
Once more, Spark is significantly faster than Dask engines and Bag is slightly
faster than Delayed. Overall, the engines do not react to changes in block size.

In Figure~\ref{fig:histo_tt_block} the total time for each function is shown.
Note that at 30 blocks the workers' resources are not fully used since the
workers can process up to 64 tasks in parallel. Moreover, the overhead for
30 blocks is erroneous due to a lower amount of block than available threads.
This leads to thread based engines to have higher overhead. Given that, 
Spark has a much lower compute time which makes it significantly faster
than Dask engines. Overall, for all engines, lower block size results in lower
IO time but increases the overhead time. Again, this is due to large overhead
desynchronizing task thus reducing the NFS bottleneck.

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
        \caption{Histogram (Python) makespan}\label{fig:histo_ms_block}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
        \caption{Histogram (Python) total time}\label{fig:histo_tt_block}
    \end{subfigure}
    \caption{8 workers, 8 CPUs/worker}
\end{figure}
% \TG{We decided to omit Dask Futures as it does not have any value over Dask
% Delayed in this application.}

\subsection{Histogram (Python): Gantt chart}
Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline
Histogram experiment implemented in Python. Spark overhead is distributed
through all the execution while Bag overhead happens mostly when a thread
reads subsequent blocks and Delayed overhead is concentrated after reading
all the blocks.

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/spark_histo_gantt.png}
        \caption{Spark}\label{fig:histo_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_histo_gantt.png}%
        \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_histo_gantt.png}%
        \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
    \end{subfigure}
    \caption{Histogram (Python): 125 blocks, 8 workers, 8
    CPUs/worker}\label{fig:histo_gantt}
\end{figure}

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
Overall, the makespan for all engines does not change considerably with an increase
in the number of workers. This is likely a result of using a non-distributed
shared file system. There is no
substantial difference between all engines although Delayed is slightly faster.
Figure~\ref{fig:histo_np_tt_worker} shows the total execution time
breakdown. The overhead and read time increase proportionally to the number
of workers thus increasing the total time.

@@

In Figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all
engines. However, when computation starts the subsequent read tasks are much
shorter for Delayed. This is due the desynchronization of the task reducing the
IO bottleneck.

\begin{figure}[!tb]
    \centering
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/spark_histo_np_gantt.png}
        \caption{Spark}\label{fig:histo_np_spark_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/bag_histo_np_gantt.png}%
        \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{\columnwidth}
        \includegraphics[clip,width=\columnwidth,
        height=0.15\textheight]{images/delayed_histo_np_gantt.png}%
        \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
    \end{subfigure}
    \caption{Histogram (NumPy): 125 blocks, 8 workers, 8
    CPUs/worker}\label{fig:histo_np_gantt}
\end{figure}

%%% BIDS EXAMPLE %%%
\subsection{BidsApp example: Number of workers}
Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when
varying the amount of workers. Overall, there is no substantial difference
between the engines. The makespan scales well with the increase in workers,
however, the scaling reduces slightly when reaching 8 workers. This is due
to an increase in I/O time caused by parallel accesses to 
the shared file system.

In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
shown. Each engine has a similar total time independently of its number of workers.
Consistently with previous observations, the overhead increases substantially with the 
number of workers.

% The overhead increases
% proportionally with the number of workers (see regression lines below). This is due
% to more communication between the scheduler and workers as well as intra-worker and
% inter-worker communication.

% \hspace{1cm}\(\hat{y}=223.53586\cdot{x} + 91.616964\) (Spark)

% \hspace{1cm}\(\hat{y}=182.82659\cdot{x} - 11.43834\) (Bag)

% \hspace{1cm}\(\hat{y}=172.54183\cdot + 400.69433\) (Delayed)

% \hspace{1cm}\(\hat{y}=132.41481\cdot{x} + 90.32813\) (Futures)
